Postural developments mediate children's visual access to social information

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
random edits

\begin{figure*}
\includegraphics[width=6in]{images/detector_samples_banner.pdf}
\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group. The last column features a false positive from OpenPose and a false negative from MTCNN.}
\end{figure*}

@@

## Face and Hand Detection
We used three face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame. The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones. We used this as a benchmark for performance, and while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual]. We next capitalized on recent improvements in computer vision, testing the performance of two face detectors that both made use of Convolutional Neural Networks (CNNs) to extract face information. The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts belonging to an agent.

This second of the CNN-based algorithms--the pose detector--provided us (along with other parts) with the location of an agent's wrists, which we used as a good proxy for the presence of a hand. We did so first because we did not capture the entire visual field experienced by the infants/children. Thus, even though the video may only show the presence of a wrist, the caregiver's hand may be within the child's field of view. Second, we did so because hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction. For example, if a caregiver was holding a toy and presenting it to the infant, their wrists may be visible but their hand may not necessarily be (as it is occluded by the toy). Similarly, we used the nose (one of the 18 body parts detected) to operationalize face presence, as any half of a face necessarily contains a nose.

@@

### Detector evaluation
To evaluate face detector performance, we needed to construct an unbiased "gold set" of labeled frames that made sure to account for the relatively rare appearance of faces in the dataset. We thus hand-labeled two samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames. Each sample was comprised of an equal number of frames taken from each participant (i.e., video). Faces were classifed as present in a frame if at least half of the face was showing. Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 2. For wrist detections, the "gold set" was constructed in the same manner, with the exception that frames with a high density of wrists were taken by wrist detections made by OpenPose.

@@

For face detection, MTCNN outperformed OpenPose when taking into account only the composite F-score (`r round(perfFacesOverall$f[1],2)` MTCNN vs. `r round(perfFacesOverall$f[2],2)` OpenPose); thus the face presence analyses to follow all result from MTCNN detections. We found that although MTCNN and OpenPose perform comparably with the random sample, MTCNN performed better on the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN. ViolaJones performed quite poorly relative to the other detectors especially on the random sample.

@@

ggarrange(p1,p2)
```
We noted characteristic changes in infants' posture and orientation across this developmental time range. The proportion of time infants spent sitting decreased with age, and the proportion of time infants spent standing increased with age. Both 8-month-olds and 12-month-olds spent equivalent amounts of time either lying/crawling, which was markedly decreased in the 16-month-olds, who spent most of their time either sitting or standing (see Figure \ref{fig:posture}).  We also observed characteristic changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions (see Figure \ref{fig:posture}).

>>>>

\begin{figure*}
\centering
\includegraphics[width=6in]{images/detector_samples_banner.pdf}
\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group. The last column features a false positive from OpenPose and a false negative from MTCNN.}
\end{figure*}

@@

## Face and Hand Detection
We used three face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame. The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones. We used this as a benchmark for performance; while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual]. We next capitalized on recent improvements in computer vision, testing the performance of two face detectors that both made use of Convolutional Neural Networks (CNNs) to extract face information. The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts belonging to an agent.

This second of the CNN-based algorithms--the pose detector--provided us with the location of an agent's wrists, which we used as a proxy for the presence of a hand. We did so first because we did not capture the entire visual field experienced by the infants/children. Thus, even though the video may only show the presence of a wrist, the caregiver's hand may be within the child's field of view. Second, we did so because hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction. For example, if a caregiver was holding a toy and presenting it to the infant, their wrists may be visible but their hand may not necessarily be (as it is occluded by the toy). Similarly, we used an agent's nose (one of the 18 body parts detected) to operationalize face presence, as any half of a face necessarily contains a nose.

@@

### Detector evaluation
To evaluate face detector performance, we needed to construct an unbiased "gold set" of labeled frames that made sure to account for the relatively rare appearance of faces in the dataset. We thus hand-labeled two samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames. Each sample was comprised of an equal number of frames taken from each participant (i.e., video). Faces were classifed as present in a frame if at least half of the face was showing. Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 2. For wrist detections, the "gold set" was constructed in the same manner, with the exception that frames with a high density of wrists were taken by pose detections made by OpenPose.

@@

For face detection, MTCNN outperformed OpenPose when taking into account only the composite F-score (`r round(perfFacesOverall$f[1],2)` MTCNN vs. `r round(perfFacesOverall$f[2],2)` OpenPose); thus the face presence analyses to follow all result from MTCNN detections. We found that although MTCNN and OpenPose perform comparably with the random sample, MTCNN performed better on the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN. ViolaJones performed quite poorly relative to the other detectors especially with respect to the random sample.

@@

ggarrange(p1,p2)
```
We noted characteristic changes in infants' posture and orientation across this developmental time range. The proportion of time infants spent sitting decreased with age, and the proportion of time infants spent standing increased with age. Both 8 month-olds and 12 month-olds spent equivalent amounts of time either lying/crawling, which was markedly decreased in the 16-month-olds, who spent most of their time either sitting or standing (see Figure \ref{fig:posture}).  We also observed characteristic changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions (see Figure \ref{fig:posture}).
